### 2014-11-20 documentation/interactive_kefi_2007_bifurcation.Rmd
interactive bifurcation graph (parallel backend)

---
title: "Pair-approximation model"
author: "Florian Schneider"
date: "Thursday, November 20, 2014"
output: html_document
runtime: shiny
---

## original model (Kéfi et al 2007)

Running the pair approximation model of Kéfi et al 2007 TPB dynamically.

```{r, echo = FALSE}
# provides parallel backend
library(foreach)
library(doSNOW)

workstation <-  list(host = "162.38.184.118", user = "schneider",
                     rscript = "/usr/lib/R/bin/Rscript",
                     snowlib = "/usr/lib/R/library")

workerlist <- c(rep("localhost", times = 10))

cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')

registerDoSNOW(cl)

```

### plant and soil parameters
```{r, echo=FALSE}

inputPanel(
  sliderInput("m", label = "plant mortality:",
              min = 0.0, max = 1, value = 0.25, step = 0.01),
  sliderInput("c", label = "competition:",
              min = 0.0, max = 1, value = 0.2, step = 0.01),
  sliderInput("delta", label = "seed distpersal:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("f", label = "facilitation:",
              min = 0.0, max = 1, value = 0.9, step = 0.01), 
  sliderInput("r", label = "regeneration rate:",
              min = 0.0, max = 1, value = 0.01, step = 0.01), 
  sliderInput("d", label = "degradation rate:",
              min = 0.0, max = 1, value = 0.1, step = 0.01)
)

```

### initial conditions
```{r, echo=FALSE}

renderPlot({

# sets parameters of iterations

global <- list(  
  m = as.numeric(input$m), #intrinsic mortality
  d = as.numeric(input$d),    # degradation
  c = as.numeric(input$c) ,     # beta*g  
  delta = as.numeric(input$delta) ,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
  r = as.numeric(input$r), 	# regeneration rate
  f = as.numeric(input$f)		# local fascilitation
)

parameters <- list(
  b = seq(0,1,.005) #seq(0,1,.002)+0.001
)

iterations <- expand.grid(parameters)
iterations <- cbind(ID = 1:dim(iterations)[1],iterations, global)

rho_ini <- 0.8
rho <- c(rho_pm = rho_ini*0.4, # rho[1]
         rho_pp = rho_ini*0.4 , # rho[2]
         rho_mm = (1-rho_ini)*0.6, # rho[3]
         rho_p = rho_ini,  # rho[4]
         rho_m = (1-rho_ini)*0.9   # rho[5]
        )

foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
  setwd("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock")
  source("code/ode_kefi.r")

  model_parms <- as.list(iterations[iteration,])

  # running the ode-solver
  runmodel <- ode(y = rho, func = odesys, times = c(1,10000), parms = model_parms)

  # transfer into ouput and calculate missing rho values
  out <- as.data.frame(runmodel)
  out$rho_0 <- 1 - out$rho_p - out$rho_m
  out$rho_p0 <- with(out, rho_p-rho_pm-rho_pp)
  out$rho_m0 <- with(out, rho_m-rho_pm-rho_mm)
  out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)
  out$ID <- iteration

  return(out[dim(out)[1],])
  } -> output

plot(output$rho_p ~ iterations$b, type = "l", ylim = c(0,1))

rho_ini <- 0.01
rho <- c(rho_pm = rho_ini*0.4, # rho[1]
         rho_pp = rho_ini*0.4 , # rho[2]
         rho_mm = (1-rho_ini)*0.6, # rho[3]
         rho_p = rho_ini,  # rho[4]
         rho_m = (1-rho_ini)*0.9   # rho[5]
        )

foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
  setwd("C:/Users/SCHNEIDER/Documents/projects/CAS02_livestock")
  source("code/ode_kefi.r")

  model_parms <- as.list(iterations[iteration,])

  # running the ode-solver
  runmodel <- ode(y = rho, func = odesys, times = c(1,10000), parms = model_parms)

  # transfer into ouput and calculate missing rho values
  out <- as.data.frame(runmodel)
  out$rho_0 <- 1 - out$rho_p - out$rho_m
  out$rho_p0 <- with(out, rho_p-rho_pm-rho_pp)
  out$rho_m0 <- with(out, rho_m-rho_pm-rho_mm)
  out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)
  out$ID <- iteration

  return(out[dim(out)[1],])
  } -> output

lines(output$rho_p ~ iterations$b)

} )

```