### 2014-11-20 documentation/solving_ODEs.Rmd
solving ODEs: add simulation of gradient, add references

---
title: ' Solving ODEs in R'
author: "Florian Schneider"
date: "Thursday, November 20, 2014"
output:
  html_document:
    keep_md: yes
  pdf_document:
    fig_height: 5
    fig_width: 7
fontsize: 10pt
geometry: margin=1.2in
csl: ../manuscript/amnat.csl
bibliography: ../manuscript/cas02.bib
---

To evaluate a spatially explicit model, we can run the pair-approximation descibed by @kefi07 into equilibrium.

We translate the system of five equations given in @kefi07 for the pair-approximation model into `R` functions.

(25)
$$ \frac{d\rho_{+-}}{dt} = d(\rho_+ - \rho_{+-} - \rho_{++}) +  (\rho_- - \rho_{--} - \rho_{+-}) \times \left( \delta\rho_+ + \frac{z-1}{z}(1-\delta)\frac{\rho_+ - \rho_{+-}  \rho_{++}}{1 - \rho_{+}  \rho_{-}}  \right) \times (b-c\rho_+) - \rho_{+-}\left( r + \frac{f}{z}+\frac{z-1}{z}f\frac{\rho_{+-}}{\rho_-}+m \right) $$

where $z$ is the number of neighbors considered, i.e. $z = 4$.

```{r}
d_rho_pm <- function(rho, parms, z = 4) { 
  with(parms,                             
      d * (rho[4] - rho[1] - rho[2]) + 
        (rho[5] - rho[3] - rho[1]) * 
        (delta*rho[4] + (z-1)/z * (1 - delta) * 
           (rho[4] - rho[1] - rho[2])/(1 - rho[4] - rho[5]) ) *
        (b - c*rho[4]) - 
        rho[1]*( r + f/z + (z-1)/z * f * rho[1]/rho[5] + m)
   )
}

```

(26)

$$  \frac{\mathrm{d}\rho_{++}}{\mathrm{d}t} = 
  2(\rho_+ - \rho_{+-}  \rho_{++}) \times \left( \delta\rho_+ + \frac{1-\delta}{z} + \frac{z-1}{z}(1-\delta)\frac{\rho_+ - \rho_{+-}  \rho_{++}}{1 - \rho_{+}  \rho_{-}}  \right) \times
  (b-c\rho_+) - 2\rho_{++}m
  $$

```{r}
d_rho_pp <- function(rho, parms, z = 4) { 
  with(parms,                             
      2 * (rho[4] - rho[1] - rho[2]) * 
        (delta*rho[4] + (1-delta)/z +(z-1)/z * (1 - delta) * 
           (rho[4] - rho[1] - rho[2])/(1 - rho[4] - rho[5]) ) *
        (b - c*rho[4]) - 2*rho[2]*m
   )
}

```

(27)

$$  \frac{\mathrm{d}\rho_{--}}{\mathrm{d}t} = 
    2d(\rho_{-} - \rho_{--} - \rho_{+-}) - 2\rho_{--} 
    \left( r + \frac{z-1}{z} f \frac{\rho_{+-}}{\rho_{-}} \right) $$

```{r}
d_rho_mm <- function(rho, parms, z = 4) { 
  with(parms,                             
      2 * (rho[5] - rho[3] - rho[1]) - 
        2*rho[3]*(r+(z-1)/z*f*rho[1]/rho[5])
   )
}

```

(28)

$$  \frac{\mathrm{d}\rho_{+}}{\mathrm{d}t} = 
    \left( \delta\rho_+ + (1-\delta)\frac{\rho_+ - \rho_{+-} -  \rho_{++}}{1 - \rho_{+}  \rho_{-}}  \right) \times
    (b-c\rho_+)(1-\rho_+ - \rho_-) - m\rho_{+}
     $$

```{r}
d_rho_p <- function(rho, parms, z = 4) { 
  with(parms,                             
      (delta*rho[4] + (1 - delta) * 
           (rho[4] - rho[1] - rho[2])/(1 - rho[4] - rho[5]) ) *
        (b-c*rho[4])*(1-rho[4]-rho[5])-m*rho[4]
      )
}

```

(29)
$$  \frac{\mathrm{d}\rho_{-}}{\mathrm{d}t} = 
d (1-\rho_+ - \rho_-) - \left(r + f \frac{\rho_{+-}}{\rho_{-}}  \right)\rho_{-} $$

```{r}
d_rho_m <- function(rho, parms, z = 4) { 
  with(parms,                             
      d*(1-rho[4]-rho[5])-(r+f*rho[1]/rho[5])*rho[5]
      )
}

```

## solving the ODE-system

In R, the package 'deSolve' provides solvers for differential equations [@soetaert10].

```{r}
library("deSolve")
```

It requires a single function that calculates the current state of a system from a previous state of the system (entered as a vector, here `rho`) with a timestep `t`. The third parameter expects any parameters that are required inside the function.

I define a function `odesys()` that returns the results of the single equations concenated into a vector.

```{r}
odesys <- function (t, rho, parms = model_parms) {
  list(c(
    d_rho_pm(rho, parms),
    d_rho_pp(rho, parms),
    d_rho_mm(rho, parms),
    d_rho_p(rho, parms),
    d_rho_m(rho, parms)
    ))
}

```

**Important:** *Equations of ODE and function for ODE solver are saved in file `code/ode_kefi.r` for further use in simulations and applications.*

We need to define the constants, which I take from the example of @kefi07, Fig. 3 to evaluate this implementation (looks very similar, differences might originate from the starting conditions, which are not given precisely in the paper).

```{r}
# model parameters
model_parms <- list(
  delta = 0, # proportion of seeds dispersed globally
  b = 0.4, # environmental quality
  m = 0.1, # plant mortality
  f = 0.8, # effect of abiotic facilitation on regeneration
  r = 0.0, # intrinsic regeneration 
  d = 0.2, # rate of degradation
  c = 0.3 # global competition among plants 
  )
```

And the initial vegetation structure which is unambigously given by the values $\rho_{+-}$, $\rho_{++}$, $\rho_{--}$, $\rho_{+}$ and $\rho_{-}$. We define the starting values as a static function of $\rho_{+}$ and store all of them in a single vector called `rho` which is expected by the functions defined above.

```{r}
# initial vegetation cover
rho_ini <- 0.1

rho <- c(rho_pm = rho_ini*0.4, # rho[1]
         rho_pp = rho_ini*0.4 , # rho[2]
         rho_mm = (1-rho_ini)*0.6, # rho[3]
         rho_p = rho_ini,  # rho[4]
         rho_m = (1-rho_ini)*0.9   # rho[5]
        )
```

Finally, the model is run by calling the function `ode()`. The output is stored in object `out` and the missing $\rho$s are completed.

Note that I am determining the timesteps exponentially, to have a decline in resolution as the dynamics steady out.

```{r}
# running the ode-solver
runmodel <- ode(y = rho, func = odesys, times = 1.1^(1:53), parms = model_parms)

# transfer into ouput and calculate missing rho values
out <- as.data.frame(runmodel)
out$rho_0 <- 1 - out$rho_p - out$rho_m
out$rho_p0 <- with(out, rho_p-rho_pm-rho_pp)
out$rho_m0 <- with(out, rho_m-rho_pm-rho_mm)
out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)
```
```{r, "fig1"}
# plot
plot(rho_p ~ time, data = out, type = "l", lwd = 2, ylab = "vegetation cover", ylim = c(0,1))
lines(rho_m ~ time, data = out, lty = 3)
lines(rho_0 ~ time, data = out, col = "grey50")
legend(60,1, c("rho_+", "rho_-", "rho_0"), col = c("black", "black", "grey50"), lty = c(1,3,1), lwd = c(2,1,1), bty = "n", horiz = TRUE)
```

## simulating gradients

For the project, we need a prediction of growth for a given value of $\rho_+$. To achieve this, I am simulating a gradient of replicates with different environmental conditions.

First, we need to generate a parallel backend for the use of the `foreach()` function.

```{r, }
library(foreach)
library(doSNOW)

workstation <-  list(host = "162.38.184.118", user = "schneider",
                     rscript = "/usr/lib/R/bin/Rscript",
                     snowlib = "/usr/lib/R/library")

workerlist <- c(rep("localhost", times = 10))

cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')

registerDoSNOW(cl)

```

Then, build a database of parameters, using a global dataset (valid for all iterations) plus a gradient over $b$, stored in `parameters` (I use `expand.grid()` to blow up the parameter space if varying multiple parameters at a time).

```{r}

first_ID = 1  #

global <- list(  
  m = 0.2, #intrinsic mortality
  d = 0.1,    # degradation
  c = 0.2, 		# beta*g  
  delta = 0.9,  	# seeds dispersed; (1-del) seeds on nearest neighbourhood
  r = 0.01, 	# regeneration rate
  f = 0.9		# local fascilitation
)

parameters <- list(
  b = seq(0,1,.0025) #seq(0,1,.002)+0.001
)

iterations <- expand.grid(parameters)
iterations <- cbind(ID = 1:dim(iterations)[1],iterations, global)

```

In a first simulation run, the initial vegetation cover should be high.

```{r}
# initial vegetation cover
rho_ini <- 0.8

rho <- c(rho_pm = rho_ini*0.4, # rho[1]
         rho_pp = rho_ini*0.4 , # rho[2]
         rho_mm = (1-rho_ini)*0.6, # rho[3]
         rho_p = rho_ini,  # rho[4]
         rho_m = (1-rho_ini)*0.9   # rho[5]
        )
```

Then, call `foreach()` to run all iterations in parallel.

```{r}
foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
 model_parms <- as.list(iterations[iteration,])

  # running the ode-solver
  runmodel <- ode(y = rho, func = odesys, times = c(1,10000), parms = model_parms)

  # transfer into ouput and calculate missing rho values
  out <- as.data.frame(runmodel)
  out$rho_0 <- 1 - out$rho_p - out$rho_m
  out$rho_p0 <- with(out, rho_p-rho_pm-rho_pp)
  out$rho_m0 <- with(out, rho_m-rho_pm-rho_mm)
  out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)
  out$ID <- iteration

  return(out[dim(out)[1],])
  } -> output_high

```

In a second run, we initialise the same set of parameters, but starting from low cover:

```{r}
# initial vegetation cover
rho_ini <- 0.01

rho <- c(rho_pm = rho_ini*0.4, # rho[1]
         rho_pp = rho_ini*0.4 , # rho[2]
         rho_mm = (1-rho_ini)*0.6, # rho[3]
         rho_p = rho_ini,  # rho[4]
         rho_m = (1-rho_ini)*0.9   # rho[5]
        )

foreach(iteration = iterations$ID, .combine = rbind, .packages = c("deSolve")) %dopar% { 
 model_parms <- as.list(iterations[iteration,])

  # running the ode-solver
  runmodel <- ode(y = rho, func = odesys, times = c(1,10000), parms = model_parms)

  # transfer into ouput and calculate missing rho values
  out <- as.data.frame(runmodel)
  out$rho_0 <- 1 - out$rho_p - out$rho_m
  out$rho_p0 <- with(out, rho_p-rho_pm-rho_pp)
  out$rho_m0 <- with(out, rho_m-rho_pm-rho_mm)
  out$rho_00 <- with(out, rho_0-rho_m0-rho_p0)
  out$ID <- iteration

  return(out[dim(out)[1],])
  } -> output_low
```

```{r, "fig2"}
plot(output_high$rho_p ~ iterations$b, type = "p", pch = 20, cex =0.75, ylim = c(0,1), xlab = "environmental quality, b", ylab = "vegetation cover, rho_+")
points(output_low$rho_p ~ iterations$b, pch = 20, cex =0.75)
```

Don't forget to shut down the parallel backend.
```{r}
stopCluster(cl)
```

# References