### 2015-01-29 documentation/interactive_nullmodel.Rmd
add interactive Version of Nullmodel in R Shiny

---
title: "null-model ode"
author: "Florian Schneider"
date: "Thursday, November 20, 2014"
output: html_document
runtime: shiny
---

## original model (Kéfi et al 2007)

Running the pair approximation model of Kéfi et al 2007 TPB dynamically.

```{r, echo=FALSE}

library("deSolve")

rho <- seq(0,1,length = 100)

```

### plant and soil parameters
```{r, echo=FALSE}

inputPanel(
  sliderInput("m", label = "intrinsic mortality:",
              min = 0.0, max = 1, value = 0.05, step = 0.01),
  sliderInput("r", label = "max. growth rate:",
              min = 0.0, max = 3, value = 1, step = 0.01),
  sliderInput("b", label = "environmental quality:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("K", label = "carrying capacity:",
              min = 0.0, max = 1, value = 0.9, step = 0.01),
  sliderInput("a", label = "search efficiency:",
              min = 0.0, max = 1, value = 0.6, step = 0.01), 
  sliderInput("h", label = "handling time:",
              min = 0.0, max = 1000, value = 100, step = 1), 
  sliderInput("L", label = "livestock units:",
              min = 0.0, max = 100, value = 10, step = 1)
)

```

### initial conditions
```{r, echo=FALSE}

inputPanel(
  sliderInput("rho_1", label = "initial plant cover:",
              min = 0.0, max = 1, value = 0.9, step = 0.01)
  )

renderPlot({

 parms <- list()
  parms$m <- as.numeric(input$m)
  parms$r <- as.numeric(input$r)
  parms$b <- as.numeric(input$b) 
  parms$K <- as.numeric(input$K) 
  parms$a <- as.numeric(input$a) 
  parms$h <- as.numeric(input$h) 
  parms$L <- as.numeric(input$L)

  plot(rho,  with(parms,  m * rho + (a * rho * L)/(1 + a * h * rho)), ylab = "growth", type = "l", ylim = c(0,0.5))
  lines(rho, with(parms, r* b* rho * (1 - rho/K) ), col = "green3" )

  })

renderPlot({

d_rho <- function(rho, parms, z = 4) { 
  with(parms, r* b* rho * (1 - rho/K) - m * rho - (a * rho * L)/(1 + a * h * rho) )  
}

odesys <- function (t, rho, parms = model_parms) {
  list( d_rho(rho, parms)  )
}

 parms <- list()
  parms$m <- as.numeric(input$m)
  parms$r <- as.numeric(input$r)
  parms$b <- as.numeric(input$b) 
  parms$K <- as.numeric(input$K) 
  parms$a <- as.numeric(input$a) 
  parms$h <- as.numeric(input$h) 
  parms$L <- as.numeric(input$L)

# running the ode-solver
runmodel <- ode(y = input$rho_1, func = odesys, times = 1.1^(1:53), parms = parms)

# transfer into ouput and calculate missing rho values
out <- as.data.frame(runmodel)

plot(out, type = "l", ylim = c(0,1))
axis(2, at = round(tail(out,1)[["1"]],2), las = 1 )

} )
```