### 2015-02-04 documentation/analytical.Rmd
work on analytical solution, simplifications

---
title: "Untitled"
author: "Florian Schneider"
date: "Tuesday, February 03, 2015"
output: html_document
---

Consider the equation system:

$$  \frac{\mathrm{d}\rho_{11} }{\mathrm{d}t} = 2 (\rho_1 - \rho_{11}) \frac{r \rho_1^{1+\alpha} (b + (1 - b)  f \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a  + v \frac{\rho_{11}}{\rho_1})L (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{q}}{1+(a  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}  $$

$$  \frac{\mathrm{d}\rho_{1}}{\mathrm{d}t} = r \rho_1^{1+\alpha} (b + (1 - b) f  \frac{\rho_{1} - \rho_{11}}{1 - \rho_1}) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) - \rho_{1}m -  \frac{(a  + v \frac{\rho_{11}}{\rho_1})L (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{1+q}}{1+(a  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}   $$

assume:

$$  \frac{\mathrm{d}\rho_{11} }{\mathrm{d}t} = 0 $$

then,

$$  0 = 2 (\rho_1 - \rho_{11}) \frac{r \rho_1^{1+\alpha} (b + (1 - b)  f \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})})}{1-\rho_1} - 2 \rho_{11} m -  2 \rho_{11} \frac{(a  + v \frac{\rho_{11}}{\rho_1})L (1 - p \frac{\rho_{11}}{\rho_1})\rho_1^{q}}{1+(a  + v \frac{\rho_{11}}{\rho_1})h\rho_1^{1+q}}  $$

$$  \rho_{11} m +  \rho_{11} \frac{(a  + v \rho_{11}\rho_1^{-1})L (1 - p \rho_{11}\rho_1^{-1})\rho_1^{q}}{1+(a  + v\rho_{11}\rho_1^{-1})h\rho_1^{1+q}} =  (\rho_1 - \rho_{11}) r \rho_1^{1+\alpha} (b + (1 - b)  f \frac{\rho_{1} - \rho_{11}}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) \frac{1}{1-\rho_1}   $$

$$  \rho_{11} m +  \frac{(a\rho_1^{q} \rho_{11}  + v \rho_{11}^2\rho_1^{q-1}) (L - L p \rho_{11}\rho_1^{-1})}{1+(a  + v\rho_{11}\rho_1^{-1})h\rho_1^{1+q}} =  (\rho_1 - \rho_{11}) r \rho_1^{1+\alpha} (b + (f - bf) (\rho_{1} - \rho_{11}) \frac{1}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) \frac{1}{1-\rho_1}   $$

$$  \rho_{11} m +  \frac{aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  }{1+(a  + v\rho_{11}\rho_1^{-1})h\rho_1^{1+q}} =  (\rho_1 - \rho_{11}) r \rho_1^{1+\alpha} (b + (f - bf) (\rho_{1} - \rho_{11}) \frac{1}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) \frac{1}{1-\rho_1}   $$

$$  \rho_{11} m +  \frac{aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  }{1+ah\rho_1^{1+q}  + vh\rho_1^{q}\rho_{11}} =  (\rho_1 - \rho_{11}) r \rho_1^{1+\alpha} (b + (f - bf) (\rho_{1} - \rho_{11}) \frac{1}{1 - \rho_1} )) (1-\frac{\rho_1}{K (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})}) \frac{1}{1-\rho_1}   $$

$$  \rho_{11} m +  \frac{aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  }{1+ah\rho_1^{1+q}  + vh\rho_1^{q}\rho_{11}} =  (\rho_1 r \rho_1^{1+\alpha}  - \rho_{11} r \rho_1^{1+\alpha} )(b + (f\rho_{1} - f\rho_{11} - bf\rho_{1} + bf \rho_{11}) \frac{1}{1 - \rho_1} )) (1-\rho_1K^{-1} (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})^{-1}) \frac{1}{1-\rho_1}   $$

$$  \rho_{11} m +  \frac{aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  }{1+ah\rho_1^{1+q}  + vh\rho_1^{q}\rho_{11}} =  (\rho_1 r \rho_1^{1+\alpha}  - \rho_{11} r \rho_1^{1+\alpha} )(b + f\rho_{1}(1 - \rho_1)^{-1} - f\rho_{11}(1 - \rho_1)^{-1} - bf\rho_{1}(1 - \rho_1)^{-1} + bf \rho_{11}(1 - \rho_1)^{-1}  ) (1-\rho_1K^{-1} (1 - c \frac{\rho_{1} - \rho_{11}}{1 - \rho_1})^{-1}) (1 - \rho_1)^{-1}    $$

$$  \rho_{11} m +  \frac{aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  }{1+ah\rho_1^{1+q}  + vh\rho_1^{q}\rho_{11}} =  (\rho_1 r \rho_1^{1+\alpha}  - \rho_{11} r \rho_1^{1+\alpha} )(b  (1 - \rho_1)^{-1}  + f\rho_{1}(1 - \rho_1)^{-2} - f\rho_{11}(1 - \rho_1)^{-2} - bf\rho_{1}(1 - \rho_1)^{-2} + bf \rho_{11}(1 - \rho_1)^{-2}  ) (1-\rho_1K^{-1} (1 - c (\rho_{1} - \rho_{11})\frac{1}{1 - \rho_1})^{-1})  $$

$$   m\rho_{11} + amh\rho_1^{1+q}\rho_{11} + vmh\rho_1^{q}\rho_{11}^2  + aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  =  (\rho_1 r \rho_1^{1+\alpha}  - \rho_{11} r \rho_1^{1+\alpha} )(b  (1 - \rho_1)^{-1}  + f\rho_{1}(1 - \rho_1)^{-2} - f\rho_{11}(1 - \rho_1)^{-2} - bf\rho_{1}(1 - \rho_1)^{-2} + bf \rho_{11}(1 - \rho_1)^{-2}  ) (1-\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (1+ah\rho_1^{1+q}  + vh\rho_1^{q}\rho_{11}) $$

$$ m\rho_{11} + amh\rho_1^{1+q}\rho_{11} + vmh\rho_1^{q}\rho_{11}^2  + aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  = (1 - \rho_1)^{-1}(b    + f\rho_{1}(1 - \rho_1)^{-1} - f\rho_{11}(1 - \rho_1)^{-1} - bf\rho_{1}(1 - \rho_1)^{-1} + bf \rho_{11}(1 - \rho_1)^{-1}  ) (1-\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (r \rho_1^{2+\alpha}  - r \rho_1^{1+\alpha} \rho_{11} +  arh\rho_1^{3+q+\alpha} - rah\rho_1^{2+q+\alpha}\rho_{11}  + vrh\rho_1^{2+q+\alpha}\rho_{11} - vrh\rho_1^{1+q+\alpha}\rho_{11}^2   )  $$

$$ m\rho_{11} + amh\rho_1^{1+q}\rho_{11} + vmh\rho_1^{q}\rho_{11}^2  + aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2}  = (1 - \rho_1)^{-2}(b (1 - \rho_1)^{1}   + f\rho_{1} - f\rho_{11} - bf\rho_{1} + bf \rho_{11} ) (1-\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (r \rho_1^{2+\alpha}  - r \rho_1^{1+\alpha} \rho_{11} +  arh\rho_1^{3+q+\alpha} - rah\rho_1^{2+q+\alpha}\rho_{11}  + vrh\rho_1^{2+q+\alpha}\rho_{11} - vrh\rho_1^{1+q+\alpha}\rho_{11}^2   )  $$

$$ ( m\rho_{11} + amh\rho_1^{1+q}\rho_{11} + vmh\rho_1^{q}\rho_{11}^2  + aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2} )(1 - \rho_1)^{2} = (b - b\rho_1 + f\rho_{1} - f\rho_{11} - bf\rho_{1} + bf \rho_{11} ) (1-\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (r \rho_1^{2+\alpha}  - r \rho_1^{1+\alpha} \rho_{11} +  arh\rho_1^{3+q+\alpha} - rah\rho_1^{2+q+\alpha}\rho_{11}  + vrh\rho_1^{2+q+\alpha}\rho_{11} - vrh\rho_1^{1+q+\alpha}\rho_{11}^2   ) $$

$$ ( m\rho_{11} + amh\rho_1^{1+q}\rho_{11} + vmh\rho_1^{q}\rho_{11}^2  + aL\rho_1^{q} \rho_{11} - aLp \rho_{11}^2\rho_1^{q-1} + v L  \rho_{11}^2\rho_1^{q-1} - v L p \rho_{11}^3\rho_1^{q-2} )(1 - \rho_1)^{2} = (b - (b+f-bf) \rho_1 - (f +bf)\rho_{11} ) (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (r \rho_1^{2+\alpha}  - r \rho_1^{1+\alpha}\rho_{11} + arh\rho_1^{3+q+\alpha} - rah\rho_1^{2+q+\alpha}\rho_{11}  + vrh\rho_1^{2+q+\alpha}\rho_{11} - vrh\rho_1^{1+q+\alpha}\rho_{11}^2   ) $$

to simplify, I assume $q = 0$ and $\alpha = 0$:

$$ ( m\rho_{11} + amh\rho_1\rho_{11} + aL\rho_1\rho_{11} + vmh\rho_1\rho_{11}^2   - aLp \rho_1^{-1}\rho_{11}^2 + v L\rho_1^{-1}  \rho_{11}^2 - v L p \rho_1^{-2}\rho_{11}^3 )(1 - \rho_1)^{2} = (b - (b+f-bf) \rho_1 - (f +bf)\rho_{11} ) (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) (  r \rho_1^{2} + arh\rho_1^{3} - r \rho_1^{1}\rho_{11}  - rah\rho_1^{2}\rho_{11}  + vrh\rho_1^{2}\rho_{11} - vrh\rho_1\rho_{11}^2   ) $$

$$ ( m\rho_{11} + amh\rho_1\rho_{11} + aL\rho_1\rho_{11} + vmh\rho_1\rho_{11}^2   - aLp \rho_1^{-1}\rho_{11}^2 + v L\rho_1^{-1}  \rho_{11}^2 - v L p \rho_1^{-2}\rho_{11}^3 )(1 - \rho_1)^{2} =  (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) ( br \rho_1^{2} - (br+fr-brf) \rho_1^{3} + barh\rho_1^{3}- (barh+farh-bfarh) \rho_1^{4} - (fr +brf) \rho_1^{2}\rho_{11} - (farh +bfarh)\rho_1^{3}\rho_{11}  - br \rho_1^{1}\rho_{11} + (br+fr-bfr) \rho_1^{2}\rho_{11} + (fr +brf) \rho_1^{1}\rho_{11}^2  - brah\rho_1^{2}\rho_{11} + (brah+frah-bfrah) \rho_1^{3}\rho_{11} + (rahf +bfrah)\rho_1^{2}\rho_{11}^2 + bvrh\rho_1^{2}\rho_{11} - (bvrh+fvrh-bfvrh) \rho_1^{3}\rho_{11} - (fvrh +bfvrh)\rho_1^{2}\rho_{11}^2  - bvrh\rho_1\rho_{11}^2 + (bvrh+fvrh-bvrhf) \rho_1^2\rho_{11}^2 + (fvrh +bvrhf)\rho_1\rho_{11}^3 ) $$

$$ ( m\rho_{11} + amh\rho_1\rho_{11} + aL\rho_1\rho_{11} + vmh\rho_1\rho_{11}^2   - aLp \rho_1^{-1}\rho_{11}^2 + v L\rho_1^{-1}  \rho_{11}^2 - v L p \rho_1^{-2}\rho_{11}^3 )(1 - \rho_1)^{2} =  (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) ( br \rho_1^{2} - (br+fr-brf+ barh) \rho_1^{3} - (barh+farh-bfarh) \rho_1^{4} - br \rho_1\rho_{11}- (2fr+br-brah+ bvrh) \rho_1^{2}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{3}\rho_{11}+ (fr +brf- bvrh) \rho_1\rho_{11}^2  + (rahf +bfrah+ bvrh)\rho_1^{2}\rho_{11}^2 + (fvrh +bvrhf)\rho_1\rho_{11}^3 ) $$

$$  m\rho_{11} - (2-amh-aL)\rho_1\rho_{11} + (m- 2amh- 2aL)\rho_1^{2}\rho_{11}+ amh\rho_1^{3}\rho_{11} + aL\rho_1^{3}\rho_{11}+ (vL- aLp) \rho_1^{-1}\rho_{11}^2 + (vmh- aLp+ vL)\rho_1\rho_{11}^2- 2vmh\rho_1^2\rho_{11}^2 + vmh\rho_1^{3}\rho_{11}^2+ 2aLp\rho_{11}^2 - 2vL\rho_{11}^2 - vLp\rho_1^{-2}\rho_{11}^3 + 2vLp\rho_1^{-1}\rho_{11}^3 - vLp  \rho_{11}^3 =  (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) ( br \rho_1^{2} - (br+fr-brf+ barh) \rho_1^{3} - (barh+farh-bfarh) \rho_1^{4} - br \rho_1\rho_{11}- (2fr+br-brah+ bvrh) \rho_1^{2}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{3}\rho_{11}+ (fr +brf- bvrh) \rho_1\rho_{11}^2  + (rahf +bfrah+ bvrh)\rho_1^{2}\rho_{11}^2 + (fvrh +bvrhf)\rho_1\rho_{11}^3 ) $$

I substitute the left hand side

$$ x =  m\rho_{11} - (2-amh-aL)\rho_1\rho_{11} + (m- 2amh- 2aL)\rho_1^{2}\rho_{11}+ amh\rho_1^{3}\rho_{11} + aL\rho_1^{3}\rho_{11}+ (vL- aLp) \rho_1^{-1}\rho_{11}^2 + (vmh- aLp+ vL)\rho_1\rho_{11}^2- 2vmh\rho_1^2\rho_{11}^2 + vmh\rho_1^{3}\rho_{11}^2+ 2aLp\rho_{11}^2 - 2vL\rho_{11}^2 - vLp\rho_1^{-2}\rho_{11}^3 + 2vLp\rho_1^{-1}\rho_{11}^3 - vLp  \rho_{11}^3$$

and the right hand part

$$ y =  ( br \rho_1^{2} - (br+fr-brf+ barh) \rho_1^{3} - (barh+farh-bfarh) \rho_1^{4} - br \rho_1\rho_{11}- (2fr+br-brah+ bvrh) \rho_1^{2}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{3}\rho_{11}+ (fr +brf- bvrh) \rho_1\rho_{11}^2  + (rahf +bfrah+ bvrh)\rho_1^{2}\rho_{11}^2 + (fvrh +bvrhf)\rho_1\rho_{11}^3 )$$

to get the simplified

$$ x =  (1 -\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1}) y$$

$$ x =  y - y\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1} $$

$$ x - y = - y\rho_1K^{-1} (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  )^{-1} $$

$$ x - y (1 - (c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1})  ) = - y\rho_1K^{-1} $$

$$ (x - y) - (x - y)(c \rho_{1}(1 - \rho_1)^{-1} - c \rho_{11}(1 - \rho_1)^{-1}) = - y\rho_1K^{-1} $$

$$ (x - y) - (x - y)(c \rho_{1} - c \rho_{11})(1 - \rho_1)^{-1} = - y\rho_1K^{-1} $$

$$ x - y + y\rho_1K^{-1} =  (x - y)(c \rho_{1} - c \rho_{11})(1 - \rho_1)^{-1}$$

$$( xK - yK + y\rho_1 )(1 - \rho_1) =  (xK - yK)(c \rho_{1} - c \rho_{11})$$

$$xK - yK + y\rho_1 - xK\rho_1 + yK\rho_1 - y\rho_1^2 =  (xK - yK)(c \rho_{1} - c \rho_{11})$$

$$xK - yK + y\rho_1 - xK\rho_1 + yK\rho_1 - y\rho_1^2 =  xKc\rho_{1} - yKc\rho_{1} -xK c \rho_{11} + yK c \rho_{11}$$

Now I resubstitute $x$ and $y$:

$$xK - yK + y\rho_1 - xK\rho_1 + yK\rho_1 - y\rho_1^2 =  xKc\rho_{1} - yKc\rho_{1} - xKc\rho_{11} + yKc \rho_{11}$$

Which gives the messy:

$$mK \rho_{11} - (2-amh-aL)K \rho_1\rho_{11} + (m- 2amh- 2aL)K \rho_1^{2}\rho_{11}+ amhK \rho_1^{3}\rho_{11} + aLK \rho_1^{3}\rho_{11}+ (vL- aLp)K  \rho_1^{-1}\rho_{11}^2 + (vmh- aLp+ vL)K \rho_1\rho_{11}^2- 2vmhK \rho_1^2\rho_{11}^2 + vmhK \rho_1^{3}\rho_{11}^2+ 2aLpK \rho_{11}^2 - 2vLK \rho_{11}^2 - vLpK \rho_1^{-2}\rho_{11}^3 + 2vLpK \rho_1^{-1}\rho_{11}^3 - vLpK \rho_{11}^3    - brK \rho_1^{2} + (br+fr-brf+ barh)K \rho_1^{3} + (barh+farh-bfarh)K \rho_1^{4} + brK \rho_1\rho_{11} + (2fr+br-brah+ bvrh)K \rho_1^{2}\rho_{11} + (2farh + brah-bvrh-fvrh+bfvrh)K\rho_1^{3}\rho_{11} - (fr +brf- bvrh)K \rho_1\rho_{11}^2 - (rahf +bfrah+ bvrh)K\rho_1^{2}\rho_{11}^2 - (fvrh +bvrhf)K\rho_1\rho_{11}^3     + br \rho_1^{3} - (br+fr-brf+ barh) \rho_1^{4} - (barh+farh-bfarh) \rho_1^{5} - br \rho_1^2\rho_{11}- (2fr+br-brah+ bvrh) \rho_1^{3}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{4}\rho_{11}+ (fr +brf- bvrh) \rho_1^2\rho_{11}^2  + (rahf +bfrah+ bvrh)\rho_1^{3}\rho_{11}^2 + (fvrh +bvrhf)\rho_1^2\rho_{11}^3     - mK \rho_1\rho_{11} + (2-amh-aL)K \rho_1^2\rho_{11} - (m- 2amh- 2aL)K \rho_1^{3}\rho_{11}- amhK \rho_1^{4}\rho_{11} - aLK \rho_1^{4}\rho_{11}- (vL- aLp)K  \rho_{11}^2 - (vmh- aLp+ vL)K \rho_1^2\rho_{11}^2+ 2vmhK \rho_1^3\rho_{11}^2 - vmhK \rho_1^{4}\rho_{11}^2- 2aLpK \rho_1 \rho_{11}^2 + 2vLK \rho_1\rho_{11}^2 + vLpK \rho_1^{-1}\rho_{11}^3 - 2vLpK \rho_{11}^3 + vLpK \rho_1\rho_{11}^3    + brK \rho_1^{3} - (br+fr-brf+ barh)K \rho_1^{4} - (barh+farh-bfarh)K \rho_1^{5} - brK \rho_1^2\rho_{11}- (2fr+br-brah+ bvrh)K \rho_1^{3}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)K\rho_1^{4}\rho_{11}+ (fr +brf- bvrh)K \rho_1^2\rho_{11}^2  + (rahf +bfrah+ bvrh)K\rho_1^{3}\rho_{11}^2 + (fvrh +bvrhf)K\rho_1^2\rho_{11}^3     - br \rho_1^{4} + (br+fr-brf+ barh) \rho_1^{5} + (barh+farh-bfarh) \rho_1^{6} + br \rho_1^3\rho_{11}+ (2fr+br-brah+ bvrh) \rho_1^{4}\rho_{11} + (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{5}\rho_{11}- (fr +brf- bvrh) \rho_1^3\rho_{11}^2  - (rahf +bfrah+ bvrh)\rho_1^{4}\rho_{11}^2 - (fvrh +bvrhf)\rho_1^3\rho_{11}^3     + brKc \rho_1^{3} - (br+fr-brf+ barh)Kc \rho_1^{4} - (barh+farh-bfarh)Kc \rho_1^{5} - brKc \rho_1^2\rho_{11}- (2fr+br-brah+ bvrh)Kc \rho_1^{3}\rho_{11} - (2farh + brah-bvrh-fvrh+bfvrh)Kc\rho_1^{4}\rho_{11}+ (fr +brf- bvrh)Kc \rho_1^2\rho_{11}^2  + (rahf +bfrah+ bvrh)Kc\rho_1^{3}\rho_{11}^2 + (fvrh +bvrhf)Kc\rho_1^2\rho_{11}^3     - mKc \rho_1\rho_{11} + (2-amh-aL)Kc \rho_1^2\rho_{11} - (m- 2amh- 2aL)Kc \rho_1^{3}\rho_{11}- amhKc \rho_1^{4}\rho_{11} - aLKc \rho_1^{4}\rho_{11}- (vL- aLp)Kc \rho_{11}^2 - (vmh- aLp+ vL)Kc \rho_1^2\rho_{11}^2+ 2vmhKc \rho_1^3\rho_{11}^2 - vmhKc \rho_1^{4}\rho_{11}^2- 2aLpKc \rho_1 \rho_{11}^2 + 2vLKc \rho_1\rho_{11}^2 + vLpKc \rho_1^{-1}\rho_{11}^3 - 2vLpKc \rho_{11}^3 + vLpKc \rho_1\rho_{11}^3    + mKc \rho_{11}^2 - (2-amh-aL)Kc \rho_1\rho_{11}^2 + (m- 2amh- 2aL)Kc \rho_1^{2}\rho_{11}^2+ amhKc \rho_1^{3}\rho_{11}^2 + aLKc \rho_1^{3}\rho_{11}^2+ (vL- aLp)Kc  \rho_1^{-1}\rho_{11}^3 + (vmh- aLp+ vL)Kc \rho_1\rho_{11}^3- 2vmhKc \rho_1^2\rho_{11}^3 + vmhKc \rho_1^{3}\rho_{11}^3+ 2aLpKc \rho_{11}^3 - 2vLKc \rho_{11}^3 - vLpKc \rho_1^{-2}\rho_{11}^4 + 2vLpKc \rho_1^{-1}\rho_{11}^4 - vLpKc \rho_{11}^4    - brKc \rho_1^{2}\rho_{11} + (br+fr-brf+ barh)Kc \rho_1^{3}\rho_{11} + (barh+farh-bfarh)Kc \rho_1^{4}\rho_{11} + brKc \rho_1\rho_{11}^2 + (2fr+br-brah+ bvrh)Kc \rho_1^{2}\rho_{11}^2 + (2farh + brah-bvrh-fvrh+bfvrh)Kc\rho_1^{3}\rho_{11}^2 - (fr +brf- bvrh)Kc \rho_1\rho_{11}^3 - (rahf +bfrah+ bvrh)Kc\rho_1^{2}\rho_{11}^3 - (fvrh +bvrhf)Kc\rho_1\rho_{11}^4 = 0 $$

I think I stop here, since there are fourth order polinomials that will not resolve.

$$0 =  (br+fr-brf+ barh)K \rho_1^{3} - brK \rho_1^{2} + (barh+farh-bfarh)K \rho_1^{4}   + br \rho_1^{3} - (br+fr-brf+ barh) \rho_1^{4} - (barh+farh-bfarh) \rho_1^{5} + brK \rho_1^{3}   - (br+fr-brf+ barh)K \rho_1^{4}- (barh+farh-bfarh)K \rho_1^{5}  - br \rho_1^{4} + (br+fr-brf+ barh) \rho_1^{5} + (barh+farh-bfarh) \rho_1^{6}  + brKc \rho_1^{3} - (br+fr-brf+ barh)Kc \rho_1^{4} - (barh+farh-bfarh)Kc \rho_1^{5}+ ( mK + (m- 2amh- 2aL)K \rho_1^{2} - (2-amh-aL)K \rho_1 + amhK \rho_1^{3} + aLK \rho_1^{3} + brK \rho_1 + (2fr+br-brah+ bvrh)K \rho_1^{2} + (2farh + brah-bvrh-fvrh+bfvrh)K\rho_1^{3} - br \rho_1^2- (2fr+br-brah+ bvrh) \rho_1^{3}  - (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{4} - mK \rho_1 + (2-amh-aL)K \rho_1^2  - (m- 2amh- 2aL)K \rho_1^{3}- amhK \rho_1^{4} - aLK \rho_1^{4}- brK \rho_1^2- (2fr+br-brah+ bvrh)K \rho_1^{3} - (2farh + brah-bvrh-fvrh+bfvrh)K\rho_1^{4}+ br \rho_1^3+ (2fr+br-brah+ bvrh) \rho_1^{4} + (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{5} - brKc \rho_1^2- (2fr+br-brah+ bvrh)Kc \rho_1^{3} - (2farh + brah-bvrh-fvrh+bfvrh)Kc\rho_1^{4} - mKc \rho_1 + (2-amh-aL)Kc \rho_1^2 - (m- 2amh- 2aL)Kc \rho_1^{3}- amhKc \rho_1^{4} - aLKc \rho_1^{4} - brKc \rho_1^{2} + (br+fr-brf+ barh)Kc \rho_1^{3} + (barh+farh-bfarh)Kc \rho_1^{4} ) \rho_{11}+ ((vL- aLp)K  \rho_1^{-1} + (vmh- aLp+ vL)K \rho_1- 2vmhK \rho_1^2 + vmhK \rho_1^{3}+ 2aLpK  - 2vLK   - (fr +brf- bvrh)K \rho_1 - (rahf +bfrah+ bvrh)K\rho_1^{2} + (fr +brf- bvrh) \rho_1^2  + (rahf +bfrah+ bvrh)\rho_1^{3}- (vL- aLp)K  - (vmh- aLp+ vL)K \rho_1^2 + 2vmhK \rho_1^3  - vmhK \rho_1^{4}- 2aLpK \rho_1  + 2vLK \rho_1   + (fr +brf- bvrh)K \rho_1^2  + (rahf +bfrah+ bvrh)K\rho_1^{3}   - (fr +brf- bvrh) \rho_1^3  - (rahf +bfrah+ bvrh)\rho_1^{4}  + (fr +brf- bvrh)Kc \rho_1^2  + (rahf +bfrah+ bvrh)Kc\rho_1^{3} - (vL- aLp)Kc  - (vmh- aLp+ vL)Kc \rho_1^2+ 2vmhKc \rho_1^3 - vmhKc \rho_1^{4}- 2aLpKc \rho_1  + 2vLKc \rho_1 + mKc  - (2-amh-aL)Kc \rho_1 + (m- 2amh- 2aL)Kc \rho_1^{2}+ amhKc \rho_1^{3} + aLKc \rho_1^{3}+ brKc \rho_1 + (2fr+br-brah+ bvrh)Kc \rho_1^{2} + (2farh + brah-bvrh-fvrh+bfvrh)Kc\rho_1^{3})\rho_{11}^2- (vLpK \rho_1^{-2} + 2vLpK \rho_1^{-1} - vLpK   - (fvrh +bvrhf)K\rho_1+ (fvrh +bvrhf)\rho_1^2  + vLpK \rho_1^{-1}    - 2vLpK  + vLpK \rho_1 + (fvrh +bvrhf)K\rho_1^2  - (fvrh +bvrhf)\rho_1^3+ (fvrh +bvrhf)Kc\rho_1^2 + vLpKc \rho_1^{-1} - 2vLpKc  + vLpKc \rho_1  + (vL- aLp)Kc  \rho_1^{-1} + (vmh- aLp+ vL)Kc \rho_1- 2vmhKc \rho_1^2 + vmhKc \rho_1^{3}+ 2aLpKc  - 2vLKc  - (fr +brf- bvrh)Kc \rho_1 - (rahf +bfrah+ bvrh)Kc\rho_1^{2} ) \rho_{11}^3- (vLpKc \rho_1^{-2} + 2vLpKc \rho_1^{-1} - vLpKc  - (fvrh +bvrhf)Kc\rho_1 ) \rho_{11}^4$$

$$0 = - brK \rho_1^{2}  + (br+2brK+rfK-brfK+barhK+brKc) \rho_1^{3}- (brKc-2br+frKc-brfKc+barhKc-brK-frK+brfK+farhK-bfarhK-fr+brf-barh) \rho_1^{4} + (br+fr-brf-barhK-farhK+bfarhK-farh+bfarh-barhKc-farhKc+bfarhKc) \rho_1^{5}+ (barh+farh-bfarh) \rho_1^{6}        + (mK + (brK-mK-mKc-2K+amhK+aLK) \rho_1 + (mK- br- 2amKh- 2aLK+2frK+brK-brahK+ bvrhK+2K-amhK-aLK- brK- 2brKc+ 2Kc-amhKc-aLKc) \rho_1^{2}   + (-mK-mKc+aLK-br++3amhK2farhK+2brahK-2bvrhK-fvrhK+bfvrhK-2fr+brah-bvrh-2frK+2aLK-2frKc+brahKc-bvrhKc+2amhKc+2aLKc+frKc-brfKc+barhKc) \rho_1^{3}-  (2farh-fvrh+bfvrh-amhK-aLK-2farhK-brahK+bvrhK+fvrhK-bfvrhK+2fr+br-amhKc-aLKc-bfarhKc-farhKc +bvrhKc+fvrhKc-bfvrhKc )\rho_1^{4} + (2farh + brah-bvrh-fvrh+bfvrh)\rho_1^{5}  )     \rho_{11}

+ ((vLK- aLpK)\rho_1^{-1}+ 2aLpK  - 2vLK - vLK+ aLpK - vLKc+ aLpKc + mKc+ (vmhK+vLK-frK-brfK+bvrhK-3aLpK+2vLK-2aLpKc+2vLKc-2Kc+amhKc+aLKc+brKc) \rho_1  + (2frKc+frbrf-bvrh+brKc-brahKc+ bvrhKc- 2vmhK+frK+brfK-bvrhK- rahfK -bfrahK-bvrhK- vmhK+ aLpK- vLK + frKc+brfKc- bvrhKc +mKc- 2amhKc- 2aLKc-vmhKc+aLpKc- vLKc) \rho_1^{2} + (3farhKc + brahKc-fvrhKc+bfvrhKc+ vmhK+rahf +bfrah+ bvrh+ 2vmhK+rahfK +bfrahK+ bvrhK-fr -brf+bvrh +bfrahKc+ 2vmhKc+ amhKc + aLKc )\rho_1^{3} + (-vmhK-rahf-bfrah-bvrh- vmhKc) \rho_1^{4}    )   \rho_{11}^2

- (+vLpK \rho_1^{-2} + (2vLpK + vLpK  + vLpKc+ vLKc- aLpKc)\rho_1^{-1} - vLpK - 2vLpK + 2aLpKc  - 2vLKc  - 2vLpKc - (fvrhK +bvrhfK+ vLpK+ vLpKc + vmhKc- aLpKc+ vLKc-frKc  -brfKc + bvrhKc )\rho_1 + (fvrh +bvrhf+fvrhK +bvrhfK+fvrhKc +bvrhfKc- 2vmhKc-rahfKc -bfrahKc- bvrhKc )\rho_1^2 - (fvrh +bvrhf+ vmhKc)\rho_1^3 )   \rho_{11}^3  - (vLpKc \rho_1^{-2} + 2vLpKc \rho_1^{-1} - vLpKc  - (fvrhKc +bvrhfKc)\rho_1 )     \rho_{11}^4    $$

assuming competition $c = 0$ and attraction $v = 0$ yields a much simpler model

$$0 = - brK \rho_1^{2}  + (br+2brK+rfK-brfK+barhK) \rho_1^{3}- (-2br-brK-frK+brfK+farhK-bfarhK-fr+brf-barh) \rho_1^{4} + (br+fr-brf-barhK-farhK+bfarhK-farh+bfarh) \rho_1^{5}+ (barh+farh-bfarh) \rho_1^{6}        + (mK + (brK-mK-2K+amhK+aLK) \rho_1 + (mK- br- 2amKh- 2aLK+2frK+brK-brahK+2K-amhK-aLK- brK) \rho_1^{2}   + (-mK+aLK-br+3amhK+2farhK+2brahK-2fr+brah-2frK+2aLK) \rho_1^{3}-  (2farh-amhK-aLK-2farhK-brahK+2fr+br)\rho_1^{4} + (2farh + brah)\rho_1^{5}  )     \rho_{11}  + (aLpK\rho_1^{-1}+ 2aLpK  + aLpK + (-frK-brfK-3aLpK+2vLK) \rho_1  + (frbrf+frK+brfK-rahfK-bfrahK+aLpK) \rho_1^{2} + (rahf +bfrah+rahfK +bfrahK-fr -brf)\rho_1^{3} + (-rahf-bfrah) \rho_1^{4}    )   \rho_{11}^2   $$