### 2015-03-03 documentation/interactive_spexmodel_bifurcations.Rmd
interactive bifurcation diagram (Shiny): refer to simfunctions.r

---
title: "global-model (ODE)"
author: "Florian Schneider"
date: "Thursday, January 30, 2015"
output: html_document
runtime: shiny
---

## developping the null model

Non-spatial global model solved as ordinary differential equations. It includes global accelerating and decelerating feedbacks on mortality and grazing.

```{r, echo=FALSE}

# provides parallel backend
library(foreach)
library(doSNOW)

workstation <-  list(host = "162.38.184.118", user = "schneider",
                     rscript = "/usr/lib/R/bin/Rscript",
                     snowlib = "/usr/lib/R/library")

workerlist <- c(rep("localhost", times = 10))

cl <- makeSOCKcluster(workerlist, master="162.38.184.88", outfile='out_messages.txt')

registerDoSNOW(cl)

library("deSolve")

defparms <- list(
    m = 0.05,
    r = 1,
    b = 0.7, 
    K = 0.9, 
    a = 0.6,
    h = 50,
    L = 4, 
    alpha = 0,
    q = 0,
    c = 0,
    f = 0,
    v = 0,
    p = 0
  )
```

### growth parameters

The basic assumption is logistic plant growth, i.e. growth goes to zero as the carrying capacity is approached (decelerating effect).

As an accelerating feedback we add water runoff, which reduces growth at low cover.

```{r, echo=FALSE}

inputPanel(
  sliderInput("r", label = "max. growth rate:",
              min = 0.0, max = 3, value = defparms$r, step = 0.01),
  sliderInput("b", label = "environmental quality:",
              min = 0.0, max = 1, value =   defparms$b, step = 0.01),
  sliderInput("K", label = "carrying capacity:",
              min = 0.0, max = 1, value =  defparms$K, step = 0.01),
  sliderInput("alpha", label = "water runoff:",
              min = 0.0, max = 1, value =  defparms$alpha, step = 0.01),
  sliderInput("f", label = "local facilitation:",
              min = 0.0, max = 1, value =  defparms$f, step = 0.01),
  sliderInput("c", label = "local competition:",
              min = 0.0, max = 1, value =  defparms$c, step = 0.01)
)

```

### mortality parameters

The basic assumption is a constant intrinsic mortality of plants and a type II functional response, i.e. grazing mortality saturates with cover (decelerating effect).

As an additional accelerating effect we assume the search efficiency to be lowest at low cover and increase with cover, turning the function into a type III functional response.

```{r, echo=FALSE}

inputPanel(
  sliderInput("m", label = "intrinsic mortality:",
              min = 0.0, max = 1, value =  defparms$m, step = 0.01),
  sliderInput("a", label = "search efficiency:",
              min = 0.0, max = 1, value =  defparms$a, step = 0.01), 
  sliderInput("h", label = "handling time:",
              min = 0.0, max = 200, value =  defparms$h, step = 1), 
  sliderInput("L", label = "livestock units:",
              min = 0.0, max = 100, value =  defparms$L, step = 1),
  sliderInput("q", label = "hill coefficient:",
              min = 0.0, max = 1, value =  defparms$q, step = 0.01),
  sliderInput("v", label = "local attractivity:",
              min = 0.0, max = 1, value =  defparms$v, step = 0.01),
  sliderInput("p", label = "local protection:",
              min = 0.0, max = 5, value =  defparms$p, step = 0.01)
)

```

### initial conditions
```{r, echo=FALSE}

inputPanel(
  selectInput("gradient", label = "select gradient:",             choices = c("b", "L"), selected = "b")
  )

renderPlot({

source("../code/simfunctions.r")

  rho <- seq(0,1,length = 100)

 parms <- list(
  m = as.numeric(input$m),
  r = as.numeric(input$r),
  b = as.numeric(input$b),
  K = as.numeric(input$K),
  a = as.numeric(input$a), 
  h = as.numeric(input$h),
  L = as.numeric(input$L),
  alpha = as.numeric(input$alpha), 
  q = as.numeric(input$q),
  f = as.numeric(input$f),
  c = as.numeric(input$c),
  v = as.numeric(input$v),
  p = as.numeric(input$p) 
  )

 par(mfrow = c(1,2))

  attractor(parms, localvals = TRUE, rho_1_ini = seq(0,1, length = 21), rho_11_ini = seq(0,1, length = 11))

 if(input$gradient == "b") bifurcation(parms, "b", c(0,1))
 if(input$gradient == "L") bifurcation(parms, "L", c(0,12))

} )
```